
Measuring hedonic and utilitarian consumption patterns has proven to be a difficult undertaking in marketing literature. However, the advent of machine learning and big data have brought about a new dimension of viewing marketing questions, such as the measurement and prediction of hedonic versus utilitarian consumer behaviour. The purpose of this paper is to consider a different approach to modeling hedonic and utilitarian behaviour by making use of text-mining, explanatory and predictive models. I use Latent Dirichlet Allocation to model product descriptions augmented with Wikipedia definitions into various topics. These topics are then used to explain and predict consumer behaviour around a proxy variable for utilitarian behaviour, namely, gift purchases. The practical implication of this work is a more relational distribution of consumer purchases than would be obtained by considering products alone, which assist in having a distribution for individual consumers over a new product dimension. After successfully modeling the topics, the results indicate that households that associate with stability purchases are more related to gift purchasing. Moreover, the results show that using a neural network for prediction yields an almost 10 percent increase in predictive capability. Finally, the interaction between predictive and explanatory statistics proves to be an interesting solution, as it highlight some features in the process of theory building. In the case of this paper, it was useful to understand which amount of topics has the highest predictive capability, and use those topics as explanatory variables.




