presentation layout:



------------------------------------------------------------------------------------------------------------------------------------------------------------------
Intro: the story. 

Lit review on marketing - topics surrounding what I'm doing - people are moving towards more fair trade purchases
Lit review on ML approaches - on the methods that I use 

What is the purpose of this paper?
What are the goals set out to achieve with this paper


------------------------------------------------------------------------------------------------------------------------------------------------------------------
Data: CEX

Describe the data structure: the CEX data and the original dataset. (UCC, dep variable, description, independent variables)

* include a figure representing the data 


------------------------------------------------------------------------------------------------------------------------------------------------------------------
Method: 

Topic model:
The data is already explained. Now explain what is done with the UCC descriptions. Explain how they are factored into the prediction and explanatory models. 

* include a figure representing how the descriptions are modeled into topics

Logistic regression:

Explain the inputs and the outputs used in the model. Also, describe the purpose of using this method. It's a benchmark, from thereon out it can also be used for explanatory analysis and prediction purposes.


NN:

Explain the inputs and outputs. Explain the NN used and explain the details of the NN. 


-----------------------------------------------------------------------------------------------------------------------------------------------------------------
Results:

Preliminary findings surrounding the dataset 
Topic model graphs: the structure of the dataset before the topics are assigned, and the dataset after the topics are assigned. Look at the first codes and allocate if according to that
 (a lot of graphs, figures etc to show what the data is telling me)



What are the results 
(present results from topic modeling - include graphs on the topics mined like a count over all the topics, or a mean/summation showing dominant topics))
(present the results from the logistic regression)



-----------------------------------------------------------------------------------------------------------------------------------------------------------------
Conclusion:

What did I do? Prediction versus explanation. Looking at how ML, in particular how topic modeling can be used to extract topics from product descriptions. 













